First Exclusive Measurement of Deep Virtual Compton Scattering off $^4$He: Toward the 3D tomography of nuclei

### 2017-02-02 eg6-letters/DVCS-coh-He4/dvcs_he4_prl.tex
update the note and the coherent paper with the latest plots
Continue editing the coherent paper

Experimentally, the coherent nuclear DVCS channel is difficult to measure because of 
its small cross section and the necessity to ensure that the target remains 
intact while emitting a hard photon ($eA \rightarrow e' A' \gamma$). Figure 
\ref{fig:diags} illustrates the hand bag diagram for the coherent DVCS on 
$^4$He. Similarly to the proton case, at large photon's 4-momentum square 
$Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer $t$ 
($= (p-p')^{2}$), the DVCS handbag diagram can be factorized into two parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and is calculable with perturbative QED, while the non-perturbative 
part is parametrized in terms of GPDs, which embed the partonic structure of 
the hadron. The GPDs depend on three variables $x$, $\xi$ and $t$, which are 
indicated in Figure \ref{fig:diags} for the DVCS process. In DVCS, $\xi$ can 
be related to the Bjorken variable $x_{B}$: $\xi\approx {{x_B}\over{2-x_B}}$, 
where $x_B=\frac{Q^2}{2M\nu}$ with the proton mass $M$ and 
$\nu=E_e-E_{e^\prime}$. One can easily access $t$ as the squared momentum 
transfered to the target, however the $x$ variable cannot be experimentally 
accessed in a DVCS reaction. Hence, we measure only complex Compton Form 
Factors (CFF) [need reference] defined as: [add FORMULA for CFFs].

The study of nuclear DVCS is still in its infancy due to the challenging 
detection of the low-energy recoil nuclei in fixed target experiments. Until 
very recently, the HERMES experiment \cite{Ellinghaus:2002zw} was the only one 
to measure DVCS off heavier nuclei such as $^4$He, N, Ne, Kr and Xe, where only 
the scattered electron and the real photon are detected. In this paper, we 
report the first exclusive measurements of the coherent DVCS channel off $^4$He 
where all products of the reaction are detected including the recoiling $^4$He 
nucleus. Following this exclusive measurement, the $^4$He CFF 
($\mathcal{H}_{A}$) will be extracted experimentally in a fully model 
independent way for the first time ever. The incoherent DVCS channel 
measurement from the same data set in in preparation and will reported in 
another publication.

In a DVCS process, the number of GPDs needed to 
parametrize the partonic structure of a hadron depends on the different 
configurations between the spin of the hadron and the helicity direction of the 
struck quark.  Therefore, the partonic structure of spin zero nuclei, such as 
$^4He$ and $^{12}C$, is parametrized by only one GPD ($H_{A}(x,\xi,t)$) at 
leading twist, while 4 GPDs arise in the nucleon case.  In this work, we have 
chose the $^4$He nucleus as our target of interest because of it's spinless 
nature and it shows a clear EMC effect \cite{JSeely}, in addition of having a 
high density and it is a well-known few-body system.

This work presents the first exclusive measurement of the beam-spin asymmetry 
of the reaction $e~^{4}He\rightarrow e'~^{4}He~\gamma$. This DVCS sensitive 
observable is measurable using a polarized lepton beam on an unpolarized target 
(U). It is convenient to use the beam-spin asymmetry as a DVCS observable 
because most of the experimental normalization and acceptance issues cancel out 
in the asymmetry ratio. It is defined in terms of the cross sections as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the DVCS differential cross 
section for a positive (negative) beam helicity. Experimentally, the DVCS 
reaction is indistinguishable from the Bethe-Heitler (BH) process, where the 
final photon is emitted either from the incoming or the outgoing leptons. The 
BH process is not sensitive to GPDs and does not carry information about the 
partonic structure of the hadronic target, and it's amplitude is calculable 
from the well-known electromagnetic form factors.

@@

The experiment, CLAS-EG6, took place in the experimental Hall-B of Jefferson 
laboratory (JLab) in 2009. JLab delivers, simultaneously, a nearly 100\% duty 
factor polarized electrons into three experimental Halls (A, B, C). The data 
were collected over three months via projecting a 6.064 GeV longitudinally 
polarized beam, (83$\%$ polarization), on a 6 atm gaseous $^4$He target.  The 
Hall-B Large Acceptance Spectrometer (CLAS) basic design \cite{CLAS_ref} was 
supplemented, during the CLAS-E1DVCS1 experimental run \cite{Girod:2007aa} in 
2005, with a specially designed electromagnetic calorimeter, Inner calorimeter 
(IC). The IC has extended the photon detection acceptance of CLAS, which is 
originally from 15$^{\circ}$ to 45$^{\circ}$, to polar angle reach as minimum 
as 4$^{\circ}$.  During the same experiment, a 5 Tesla solenoid was added 
around the target to shield the inner detectors from the low-energy M\o ller 
electrons.

At 6 GeV incident electron beam energy, the recoil $^4$He nuclei, from the 
coherent DVCS channel, have an average momentum (per charge) around 125 MeV/c, 
while the CLAS spectrometer detects charged particles with a threshold of 250 
MeV/c. In order to ensure the exclusivity of the our coherent DVCS channel, we 
built a small and light radial time projection chamber (RTPC) to detect 
recoiling nuclei down to energies of few MeVs. Figure~\ref{fig:RTPC} presents 
our cylindrical RTPC, which is 20 cm long and 15 cm diameter, surrounding the 
$^4$He gaseous target and being inside the available space inside the solenoid 
, with a 3 cm radial drift length. The detector was specifically calibrated for 
$^4$He nuclei using elastic scattering produced with a 1.2~GeV electron beam.

\begin{figure}[tb]
\includegraphics[width=7.0cm,angle=-90]{figs/RTPC.pdf}
\vspace{-1.1cm}
\caption{Left: A picture of the CLAS-EG6 RTPC before insertion into the 
   solenoid. Right: A cross section of the CLAS-EG6 RTPC perpendicular to the 
   beam direction. An illustration of a $^4$He track originating from the 
   pressurized straw target is shown along with the electrons produced in the 
drift region.}
\label{fig:RTPC}
\end{figure}

%DVCS selection
Identifying the coherent DVCS candidates is the first step of the data 
analysis. These events have one electron, one $^4$He, and at least one photon 
in the final state. Electrons were identified by passing the fiducial cuts and 
having signals in all the sub-detectors of CLAS spectrometer (drift chambers, 
Cherenkov counters, the standard CLAS electromagnetic calorimeter, and 
scintillators). $^4$He tracks were identified by passing all the geometrical, 
timing and quality cuts in the RTPC detector. In principle, with constant beam 
luminosity, target density and pressure, the event rate has to be constant over 
the experimental time. Due to the changes in the experimental conditions, such 
as changing a trigger in a detector, a slight shift in the beam position or a 
system failure somewhere, this rate changes. We minimize the effects of these 
changes on the reconstructed events by selecting the good runs. To this aim, we 
monitor the ratio between the number of the good tracks reconstructed in the 
RTPC to number of the detected good electrons in the CLAS detector as a 
function of run number \cite{eg6_note}.

Even though the DVCS reaction has only one real photon in the final state, 
events with more than one good photon are not discarded at this stage. This is 
motivated by the fact that some photons correspond to random coincidences and 
discarding these events results in losing good events. Then, events with one or 
more $\pi^{0}$ are removed from the coherent DVCS sample. After that, the most 
energetic IC photon was considered as the DVCS photon candidate. Next, a 
$Q^{2}>1~[GeV^{2}/c^{2}]$ cut is applied on the DVCS candidates in order to 
ensure that the interaction occurs at the partonic level and the applicability 
of the factorization in the DVCS handbag diagram.  Once the three final state 
particles were identified with their 3-momentum vectors, the exclusivity of the 
coherent DVCS events were ensured by applying a set of exclusive cuts, which 
are: the co-planarity angle ($\Delta \phi$), missing energy, missing mass 
squared and missing transverse momentum in the $e'^4He'\gamma X$ final state 
configuration, the missing mass squared in the $e'^4He'X$ and $e'\gamma X$ 
configurations, and finally the cone angle ($\theta$) between the measured real 
photon and the missing particle in the $e'^4He'X$ configuration.  
Figure~\ref{fig:kin-cuts} presents four of the applied exclusivity cuts, where 
3$\sigma$ cuts are applied on all the exclusive quantities except the missing 
energy, for which a [-0.45,0.5] GeV cut was adopted to reduce the background 
contribution. Finally,

%asymmetries
Even with all the previously presented exclusive cuts, the selected events are
not all true DVCS events. We identified several background contributions to the 
coherent DVCS process, in particular accidental events and exclusive Deeply 
Virtual $\pi^0$ Production (DV$\pi^0$P). The accidental events where the 
different particles come from different events are suppressed by the limited 
phase space allowed by the exclusivity cuts. We estimate the accidental events 
to represent 4.1\% of our sample. This relatively large number is due to the 
small cross section of the DVCS. We evaluated this contribution by selecting 
events passing all our cuts but with particles originating from different 
vertices. Regarding the DV$\pi^0$P, which can easily be mistaken with DVCS when 
one of the two photons of the $\pi^0$ decay is produced at low energy in the 
laboratory frame. To estimate the importance of this background, we developed 
an event generator that we calibrated to match our measured experimental yield 
of exclusive $\pi^0$. We used this generator together with a GEANT3 simulation 
of our detection system to estimate the ratio of acceptance between DV$\pi^0$P 
where the two photons are detected and those where only one photon is detected 
and would pass our DVCS selection cuts. This ratio obtained from simulation is 
then multiplied by the measured yield of DV$\pi^0$P events, indicating a 
contamination of 2 to 4\%. The study of systematics errors showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%). However added quadratically, these errors sum up to 
10\%, which remain for all bins well below the statistical errors.

@@

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/coherent-ALU.pdf}
\caption{The coherent $A_{LU}$ as a function of $\phi$ in
   $Q^{2}$(top panel), $x_{B}$ (middle panel), and $-t$ (bottom panel) bins.  
   The error bars represent the statistical and the systematic uncertainties 
   added quadratically, shown on top in green are error bars representing only 
   the statistical uncertainties. The bluish-gray bands represent the 
   systematic uncertainties, including the normalisation systematic 
uncertainties. The red curves represent fits in the form of equation 
\ref{eq:A_LU-coh}.}
\label{fig:alu}
\end{figure}

@@

%Acknowledgments

>>>>

Experimentally, the coherent nuclear DVCS channel is difficult to measure because of 
its small cross section and the necessity to ensure that the target remains 
intact while emitting a hard photon ($eA \rightarrow e' A' \gamma$). Figure 
\ref{fig:diags} illustrates the hand bag diagram for the coherent DVCS on 
$^4$He. Similarly to the proton case, at large photon's 4-momentum square 
$Q^2$ ($= -(k-k')^{2}$) and small squared momentum transfer $t$ 
($= (p-p')^{2}$), the DVCS handbag diagram can be factorized into two parts 
\cite{Freund_Collins,Ji_Osborne}. The hard part includes photons-quark 
interaction and is calculable with perturbative QED, while the non-perturbative 
part is parametrized in terms of GPDs, which embed the partonic structure of 
the hadron. The GPDs depend on three variables $x$, $\xi$ and $t$, which are 
indicated in Figure \ref{fig:diags} for the DVCS process. In DVCS, $\xi$ can 
be related to the Bjorken variable $x_{B}$: $\xi\approx {{x_B}\over{2-x_B}}$, 
where $x_B=\frac{Q^2}{2M\nu}$ with $M$ the proton mass and 
$\nu=E_e-E_{e^\prime}$. One can easily access $t$ as the squared momentum 
transfered to the target, however the $x$ variable cannot be experimentally 
accessed in a DVCS reaction. Hence, we measure only complex Compton Form 
Factors (CFF) [need reference] defined as: [add FORMULA for CFFs].

The main challenge to study nuclear DVCS is to ensure the exclusivity of the 
reaction. Until now, the only available data set of nuclear DVCS
was from the HERMES experiment \cite{Ellinghaus:2002zw}, where exclusivity is 
based on kinematic cuts on the measured scattered electron and real photon only. 
The measurement was performed on a large set of nuclei ($^4$He, N, Ne, Kr and 
Xe) but has been critisized for mixing coherent and incoherent processes 
[ref]. In this regard, the direct detection of the low-energy recoil nuclei is the best way to
garanty that the nuclei is intact and that the reaction did not occur on a 
bound nucleon. The CEBAF Large Acceptance Spectrometer (CLAS) setting  in the
Hall-B of JLab being already optimized for DVCS measurments 
[refs], we built a dedicated radial time projection chamber (RTPC) for 
the detection of the recoiling low energy nuclei in order to complete this
experimental setting. The $^4$He nucleus is an ideal experimental target in
this regard as it is light enough to be detected in such a setting, while it
is subject to significant nuclear effects \cite{JSeely} and has rather high 
density. Using a helium target leads to another important advantage, the number of GPDs 
defined for a hadron depends on its spin. The structure of a spin zero nuclei, such as 
$^4He$, is parametrized by only one chiral even GPD ($H_{A}(x,\xi,t)$) at 
leading twist, while 4 GPDs arise in the nucleon case. That significantly
simplifies the interpretation of the data and allow a model independent
extraction of the $^4$He CFF ($\mathcal{H}_{A}$) that will be presented at the
end of this letter.

The experiment E08-024 took place in the experimental Hall-B of Jefferson 
laboratory (JLab) in 2009 using the nearly 100\% duty factor longitudinally 
polarized electron beam (83$\%$ polarization) at its full energy of 6.064 
GeV. The data were collected over three months using a 6 atm gaseous $^4$He 
target placed in the center of CLAS. For DVCS experiments, the CLAS base line 
design \cite{CLAS_ref} is supplemented with an inner calorimeter (IC) and a
solenoid. The IC extends the photon detection acceptance of CLAS, which is 
originally from 15$^{\circ}$ to 45$^{\circ}$, to polar angle as low 
as 4$^{\circ}$. At these small angles the low-energy M\o{}ller 
electrons produced in the target form a very high rate background that is
suppressed by a 5 Tesla solenoid placed around the target.

Because the coherent DVCS cross section is peaked at low $-t$, the recoil 
$^4$He nuclei produced in DVCS have a low average momentum around 250 MeV/c.
The CLAS cannot detect such low energy $\alpha$ particles, so in order to 
ensure the exclusivity of the measurement, we 
built a small and light RTPC to complement CLAS. Figure~\ref{fig:RTPC} presents 
the design of this RTPC that is surrounding the 
$^4$He gaseous target and inside the solenoid magnet.
The detector was specifically calibrated for the detection of
$^4$He nuclei using elastic scattering produced with a 1.2~GeV electron beam.

\begin{figure}[tb]
\vspace{-1.1cm}
\includegraphics[width=7.0cm,angle=-90]{figs/RTPC.pdf}
\vspace{-1.1cm}
\caption{Left: A picture of the E08-024 RTPC before insertion into the 
   solenoid. Right: A cross section of the E08-024 RTPC perpendicular to the 
   beam direction. An illustration of a $^4$He track originating from the 
   pressurized straw target is shown along with the electrons produced in the 
   drift region.}
\label{fig:RTPC}
\end{figure}

%DVCS selection
To identify the coherent DVCS events, we first select events where one 
electron, one $^4$He and at least one photon are detected in the final state. 
Electrons are identified by using fiducial cuts and detecting a good
signals in all the sub-detectors of the base line CLAS (drift chambers, 
Cherenkov counters, electromagnetic calorimeter, and time of flight 
scintillators). The $^4$He tracks are identified with fiducial, 
timing and track quality cuts in the RTPC detector. In addition,
we apply a vertex matching cut to make sure the electron and the helium
nuclei originate from the same place. The photons are detected either
in the IC or the CLAS electromagnetic calorimeter, note that even though the 
DVCS reaction has only one real photon in the final state, 
events with more than one good photon are not discarded at this stage. This is 
motivated by the fact that soft photons are likely to be produced in random 
coincidences, but cannot be mistaken for DVCS photons which are always at high
energies ($>2$ GeV).

Then, events with one or 
more $\pi^{0}$ are removed from the coherent DVCS sample. After that, the most 
energetic IC photon was considered as the DVCS photon candidate. Next, a 
$Q^{2}>1~[GeV^{2}/c^{2}]$ cut is applied on the DVCS candidates in order to 
ensure that the interaction occurs at the partonic level and the applicability 
of the factorization in the DVCS handbag diagram.  Once the three final state 
particles were identified with their 3-momentum vectors, the exclusivity of the 
coherent DVCS events were ensured by applying a set of exclusive cuts, which 
are: the co-planarity angle ($\Delta \phi$), missing energy, missing mass 
squared and missing transverse momentum in the $e'^4He'\gamma X$ final state 
configuration, the missing mass squared in the $e'^4He'X$ and $e'\gamma X$ 
configurations, and finally the cone angle ($\theta$) between the measured real 
photon and the missing particle in the $e'^4He'X$ configuration.  
Figure~\ref{fig:kin-cuts} presents four of the applied exclusivity cuts, where 
3$\sigma$ cuts are applied on all the exclusive quantities except the missing 
energy, for which a [-0.45,0.5] GeV cut was adopted to reduce the background 
contribution. Finally,

Even with all the previously presented exclusive cuts, the selected events are
not all true DVCS events. We identified several background contributions to the 
coherent DVCS process, in particular accidental events and exclusive Deeply 
Virtual $\pi^0$ Production (DV$\pi^0$P). The accidental events where the 
different particles come from different events are suppressed by the limited 
phase space allowed by the exclusivity cuts. We estimate the accidental events 
to represent 4.1\% of our sample. This relatively large number is due to the 
small cross section of the DVCS. We evaluated this contribution by selecting 
events passing all our cuts but with particles originating from different 
vertices. Regarding the DV$\pi^0$P, which can easily be mistaken with DVCS when 
one of the two photons of the $\pi^0$ decay is produced at low energy in the 
laboratory frame. To estimate the importance of this background, we developed 
an event generator that we calibrated to match our measured experimental yield 
of exclusive $\pi^0$. We used this generator together with a GEANT3 simulation 
of our detection system to estimate the ratio of acceptance between DV$\pi^0$P 
where the two photons are detected and those where only one photon is detected 
and would pass our DVCS selection cuts. This ratio obtained from simulation is 
then multiplied by the measured yield of DV$\pi^0$P events, indicating a 
contamination of 2 to 4\%. The study of systematics errors showed that the main 
contributions come from the choice of the DVCS exclusivity cuts (8\%) and the 
large binning size (5.1\%). However added quadratically, these errors sum up to 
10\%, which remain for all bins well below the statistical errors.

%asymmetries
Different observables are available in DVCS, the cross sections, the BSA...
This work presents the first exclusive measurement of the beam-spin asymmetry 
of the reaction $e~^{4}He\rightarrow e'~^{4}He~\gamma$. This DVCS sensitive 
observable is measurable using a polarized lepton beam on an unpolarized target 
(U). It is convenient to use the beam-spin asymmetry as a DVCS observable 
because most of the experimental normalization and acceptance issues cancel out 
in the asymmetry ratio. It is defined in terms of the cross sections as:
  \begin{equation}
  A_{LU} = \frac{d^{5}\sigma^{+} - d^{5}\sigma^{-} }
                {d^{5}\sigma^{+} + d^{5}\sigma^{-}}.
    \label{BSA_equation}
  \end{equation}
where $d^{5}\sigma^{+}$($d^{5}\sigma^{-}$) is the DVCS differential cross 
section for a positive (negative) beam helicity. Experimentally, the DVCS 
reaction is indistinguishable from the Bethe-Heitler (BH) process, where the 
final photon is emitted either from the incoming or the outgoing leptons. The 
BH process is not sensitive to GPDs and does not carry information about the 
partonic structure of the hadronic target, and it's amplitude is calculable 
from the well-known electromagnetic form factors.

@@

@@

\begin{figure}[tb]
\includegraphics[width=8.9cm]{figs/coherent-ALU.pdf}
\caption{The coherent $A_{LU}$ as a function of $\phi$ in
   $Q^{2}$(top panel), $x_{B}$ (middle panel), and $-t$ (bottom panel) bins.  
   The error bars represent the statistical uncertainties. The bluish-gray 
   bands represent the systematic uncertainties, including the normalisation 
systematic uncertainties. The red curves represent fits in the form of equation 
\ref{eq:A_LU-coh}.}
\label{fig:alu}
\end{figure}

@@

The incoherent DVCS channel 
measurement from the same data set in in preparation and will reported in 
another publication. 
%Acknowledgments
